output: rmarkdown::github_document

Jesper Maag

jespermaag@gmail.com


# Plotting the percentage vote for each political party in Sweden per county 


Loading libraries 
```{r}
library(httr)
library(lubridate)
library(jsonlite)
library(ggplot2)
library(swemaps)
library(rkolada)
library(maps)
library(dplyr)
library(RColorBrewer)
library(htmlwidgets)
library(rvg)
library(ggiraph)
library(gganimate)
library(gridExtra)
library(grid)
library(gtable)
```

#### Getting data
Get all the election data per county and per election.
This involves trying to figure out where on the Swedish Statistical Bureau you can find information (http://api.scb.se/OV0104/v1/doris/sv/ssd/) .
For english go to http://api.scb.se/OV0104/v1/doris/en/ssd
```{r}
url <- "http://api.scb.se/OV0104/v1/doris/sv/ssd/ME/ME0104/ME0104C/ME0104T3"

SCB <- jsonlite::fromJSON(url)

SCB$variables$code


query = paste0(' { "query": [  {"code": "ContentsCode","selection": {"filter": "item", "values": [ "ME0104B7"] } }, {"code": "Tid", "selection": {"filter": "item","values": ["', paste(SCB$variables$values[[4]],collapse='" "'),'"]}},   {"code": "Region","selection": {"filter": "item", "values": ["', paste(SCB$variables$values[[1]],collapse='" "'), '"] } } ],"response": {"format": "csv"}  }')

results <- POST(url, body=query)
SCBdata <- content(results)

SCBdata
```

#### Extract county code and remove lan
```{r}

SCBdata$kod <- unlist(lapply(strsplit(SCBdata$region,split=' '), function(x) {x[1]}))
SCBdata <- SCBdata[grep('VR', SCBdata$kod, invert=T),]
      
```

#### Clean data and make interactive

```{r}  
sweden_map <- map_kn %>%
	select(leaflet_long, leaflet_lat, knkod, knnamn, lnnamn) %>%
	rename(long = leaflet_long, lat= leaflet_lat, region=knkod, subregion=knnamn) %>%
	fortify() %>%
        unique()

SCBdata <- SCBdata[SCBdata$kod%in%sweden_map$region,]

SCBmelt <- melt(SCBdata, id=c('region', 'parti mm', 'kod'))
SCBmelt[SCBmelt$value=='..',]$value <- 0
SCBmelt$value <- as.numeric(SCBmelt$value)


SCBmelt$tip <- paste0(
"<b>", sub(".*? (.+)", "\\1", SCBmelt$region), "</b> ",
"<br>", paste0(round(SCBmelt$value),'%'))
SCBmelt$variable= gsub('Andel röster \\(av giltiga röster\\) ','',SCBmelt$variable )

colnames(SCBmelt) <- c('region', 'parti', 'kod', 'year','percent', 'tip')

SCBmelt <- SCBmelt[grep('ogiltiga valsedlar|ej röstande', SCBmelt$parti, invert=T),]

```


#### Create colour for each party
```{r}
unique(SCBmelt$parti)

SCBmelt$cols <- "#129B93"
SCBmelt$cols[SCBmelt$parti %in% "Socialdemokraterna"] <- "firebrick2"
SCBmelt$cols[SCBmelt$parti %in% "Moderaterna"] <- "deepskyblue4"
SCBmelt$cols[SCBmelt$parti %in% "Centerpartiet"] <- "darkgreen"
SCBmelt$cols[SCBmelt$parti %in% "Folkpartiet"] <- "dodgerblue3"
SCBmelt$cols[SCBmelt$parti %in% "Kristdemokraterna"] <- "navy"
SCBmelt$cols[SCBmelt$parti %in% "Miljöpartiet"] <- "green3"
SCBmelt$cols[SCBmelt$parti %in% "Vänsterpartiet"] <- "red3"
SCBmelt$cols[SCBmelt$parti %in% "Sverigedemokraterna"] <- "goldenrod3"
SCBmelt$cols[SCBmelt$parti %in% "övriga partier"] <- "gray52"


```

#### Plot data 
```{r}
partiPlot <- SCBmelt[SCBmelt$year=='2014',]
plotlist <- list() # create empty list to fill with plots
for(i in 1:length(unique(partiPlot$parti)))
  local({
    i <- i
    # filter data to only contain current party
    plotdata <- partiPlot %>% filter(parti %in% unique(partiPlot$parti)[i])
    plotdata$Long <- 'all'
    plotdata$Long[as.numeric(plotdata$kod)<0563] <- 'south'
    # save plot as list element  
    plotlist[[i]] <<-    plotdata %>%
                                ggplot( aes(map_id=kod, fill=percent, frame = year))+ 
                                geom_map_interactive(data=plotdata, map=sweden_map, aes(fill=percent, 
                                tooltip=tip, map_id=kod, data_id=kod), 
                                colour="white", size=0.05)+
                                expand_limits(x = sweden_map$long, y = sweden_map$lat)+
                                theme_void() +
                                scale_fill_gradient2(low="#F4F4F4", high=unique(plotdata$cols)) +
                                theme(legend.position='none')+
                                ggtitle(unique(partiPlot$parti)[i])+
                                theme(plot.background = element_rect(fill = '#999999'))+
                                theme(line = element_blank(),plot.title = element_text(hjust = 0.5))#+
                                #facet_zoom(xlim = c(10, 18), ylim = c(55, 60), zoom.size = 2)
})
SwePol <-do.call("grid.arrange", c(plotlist, ncol=3))

title <- textGrob("2014",gp=gpar(fontsize=50))
        padding <- unit(5,"mm")

table <- gtable_add_rows(
     SwePol, 
     heights = grobHeight(title) + padding,
     pos = 0)
table <- gtable_add_grob(
    table, 
    title, 
    1, 1, 1, ncol(table))



ggiraph(code = {print(grid.draw(table))},  hover_css = "fill:red;r:3pt;" ,height_svg=9)

```

#### Create a gif of all the years possible
ggannimate did not seem to work for a TableGrob object. 
Instead I used a for loop for each year and combined the different years using Imagemagick. 
```{r}
#for (j in unique(SCBmelt$year)){
#        partiPlot <- SCBmelt[SCBmelt$year==j,]
#        plotlist <- list() # create empty list to fill with plots
#        for(i in 1:length(unique(partiPlot$parti)))
#          local({
#            i <- i
#            # filter data to only contain current party
#            plotdata <- partiPlot %>% filter(parti %in% unique(partiPlot$parti)[i])
#            plotdata$Long <- 'all'
#            plotdata$Long[as.numeric(plotdata$kod)<0563] <- 'south'
#            # save plot as list element  
#            plotlist[[i]] <<-    plotdata %>%
#                                        ggplot( aes(map_id=kod, fill=percent, frame = year))+ 
#                                        geom_map_interactive(data=plotdata, map=sweden_map, aes(fill=percent, 
#                                        tooltip=tip, map_id=kod, data_id=kod), 
#                                        colour="white", size=0.05)+
#                                        expand_limits(x = sweden_map$long, y = sweden_map$lat)+
#                                        theme_void() +
#                                        scale_fill_gradient2(low="#F4F4F4", high=unique(plotdata$cols)) +
#                                        theme(legend.position='none')+
#                                        ggtitle(unique(partiPlot$parti)[i])+
#                                        theme(plot.background = element_rect(fill = '#999999'))+
#                                        theme(line = element_blank(),plot.title = element_text(hjust = 0.5))#+
#                                        #facet_zoom(xlim = c(10, 18), ylim = c(55, 60), zoom.size = 2)
#        })
#        SwePol <-do.call("grid.arrange", c(plotlist, ncol=3))
#        
#        title <- textGrob(j,gp=gpar(fontsize=50))
#        padding <- unit(5,"mm")
#
#        table <- gtable_add_rows(
#             SwePol, 
#             heights = grobHeight(title) + padding,
#             pos = 0)
#        table <- gtable_add_grob(
#            table, 
#            title, 
#            1, 1, 1, ncol(table))
#
#        jpeg(paste(j,'.png'),height=700,width=400)
#        grid.draw(table)
#        dev.off()     
#}
```
```
#convert -delay 100 -loop 0 *.png SwedenPol.gif 
```
#